Dark Control:\\ A Unified Account of Default Mode Function\\ by Control Theory and Reinforcement Learning

### 2017-08-11 paper/paper.tex
more refs + barto ideas

The environment is generally taken as \textit{stochastic},
that is, changing in random ways. In addition, the environment is only
\textit{partially observable} in the sense that only limited aspects of the environment's
state are accessible to the agent's sensory perception.
\citep{starkweather2017dopamine,o2006making,o2006making}.
We assume that volatility of the environment
is realistic in a computational model which sets out
to explain DMN functions of the human brain.
% XXX: the following passage is already in the introduction
% Indeed, it is known that the more the external world is predictable,
% the more mental activity becomes detached from the actual sensory inputs
% \citep{antrobus1966studies, pope1978regulation},
% the more stimulus-independent thoughts occur, and
% the higher the neural activity in the DMN.
% Conversely, the more the currently executed task is unknown and unpracticed,
% the lower the DMN activity
% \citep{filler1973daydreaming, teasdale1995stimulus}.
% %
% These ``offline'' processes may however contribute to
% optimizing control of the organism.
% Such a framework of DMN funtion based on reinforcement learning
% can naturally embed human behavior
% in the tension between exploitative action with immediate gains and
% explorative action with longer-term reward schedules
% \citep{dayan2008decision}.
We argue that a functional account of the DMN based on RL
can naturally embed human behavior
in the tension between exploitative action with immediate gains and
explorative action with longer-term reward outcomes
\citep{dayan2008decision}.

@@

\subsection{Interlude: Rethinking the Action-State-Reward model of RL}
``Remarkably, Barto proposes to decouple extrinsic rewards (i.e generated by the external environment) from intrinsic reward (generated from withing the "organism")
biologically, extrinsic rewards would correspond to so-called "sensations"
while intrinsic rewards would correspond to "reward signals".

Thus there are two enviroments: the external environment, and the internal environment built-into the "organism". The organism takes highly level DECISIONS (the DMN can confortably sit here!) which act on the internal environment (this environment probably corresponds to low-level subsystems like motor, etc.) which then transform theses  into ACTIONS executed in the external environment. The external environment generates sensations which get mapped into reward signals (via the internal environment), etc.
Il y a 6 min
Guillaume Dumas (deep.introspection@gmail.com)
Yeah, good for creating Skynet :D!
(Half joking 🙃)
Guillaume • Il y a 6 min
haha
The interesting thing is that with this rethinking, the organism can have a perception-action loop even in the absence of external stimulus.
This is called curiosity.''

@@

Rational probabilistic planning can be reformulated
as a standard memory Markov process by simply expanding the
definition of the state $s$ to include experience episodes of the past.
This adds the capacity for memory, since then, the next state depends not only on the current state but also on a finite number of past states.

>>>>

The environment is generally taken as \textit{stochastic},
that is, changing in random ways. In addition, the environment is only
\textit{partially observable} in the sense that only limited aspects of the environment's
state are accessible to the agent's sensory perception~\citep{starkweather2017dopamine,o2006making,o2006making}.
We assume that volatility of the environment
is realistic in a computational model which sets out
to explain DMN functions of the human brain.
% XXX: the following passage is already in the introduction
% Indeed, it is known that the more the external world is predictable,
% the more mental activity becomes detached from the actual sensory inputs
% \citep{antrobus1966studies, pope1978regulation},
% the more stimulus-independent thoughts occur, and
% the higher the neural activity in the DMN.
% Conversely, the more the currently executed task is unknown and unpracticed,
% the lower the DMN activity
% \citep{filler1973daydreaming, teasdale1995stimulus}.
% %
% These ``offline'' processes may however contribute to
% optimizing control of the organism.
% Such a framework of DMN funtion based on reinforcement learning
% can naturally embed human behavior
% in the tension between exploitative action with immediate gains and
% explorative action with longer-term reward schedules
% \citep{dayan2008decision}.
We argue that a functional account of the DMN based on RL
can naturally embed human behavior
in the tension between exploitative action with immediate gains and
explorative action with longer-term reward outcomes
\citep{dayan2008decision}.

@@

\subsection{Interlude: Rethinking the Action-State-Reward model of RL}
Barto proposes to decouple extrinsic rewards (i.e generated by the external environment) from intrinsic reward (generated from withing the "organism")
biologically, extrinsic rewards would correspond to so-called "sensations"
while intrinsic rewards would correspond to "reward signals".

Thus there are two environments: the external environment, and the internal environment built-into the "organism". The organism takes highly level DECISIONS (the DMN can confortably sit here!) which act on the internal environment (this environment probably corresponds to low-level subsystems like motor, etc.) which then transform theses  into ACTIONS executed in the external environment. The external environment generates sensations which get mapped into reward signals (via the internal environment), etc.

The interesting thing is that with this rethinking, the organism can have a perception-action loop even in the absence of external stimulus.
This is called curiosity...

@@

Rational probabilistic planning can be reformulated
as a standard memory Markov process by simply expanding the
definition of the state $s$ to include experience episodes of the past.
This adds the capacity for memory, since then, the next state depends not only on the current state but also on a finite number of past states. This is precisely what partially observable MDPs do ~\citep{starkweather2017dopamine,o2006making,o2006making}.
